---
title: "R_607_Project_2_Wide_to_Long_Transform_Data_Daniel_Thonn"
output: html_document
course: R-607 CUNY MSDA program

---

## Assignment 607_Project_2_Wide_to_Long_Transform_Birth Data_Daniel_Thonn

Summary of Assignment
This assignment involves Tidying and Transforming data

This Assignment requires the following:

1). R-Studio

The following R-packages are used:
1.tidyr
2.dplyr
3.ggplot2 

Steps to reproduce:

1). Place the file locally: C:/mysqldata/Wide_Birth_Data_2.csv

2). run the R-Studio file: R_607_Project_2_Wide_to_Long_Daniel_Thonn_3a.Rmd


Setting up and Preparing the Environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}

#install.packages("stringr")
#install.packages("tidyr")
#install.packages("dplyr")
#install.packages("plyr")
#library(stringr)
library(tidyr)
library(dplyr)
#library(plyr)
#detach("package:plyr", unload=TRUE)
library(ggplot2)

```


Load the character names data into an R dataframe, cleanup, and convert to long format


```{r, echo=TRUE}

# Load the data csv file to a vector
d <- read.csv(
              "C:/mysqldata/Wide_Birth_Data_2.csv",
              sep=",",
              na.strings = "",
              blank.lines.skip = TRUE,
                stringsAsFactors=FALSE)
list(d)

# Convert the input vector to a dataframe
df_Birth_Data_1 = data.frame(d)
df_Birth_Data_1

df_Birth_Data_1b <- df_Birth_Data_1
df_Birth_Data_1b

names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "v012"] <- "MothersAge"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b2_01"] <- "BirthYear1"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b2_02"] <- "Birthyear2"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b2_03"] <- "Birthyear3"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b4_01"] <- "Sexbirth1"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b4_02"] <- "Sexbirth2"
names(df_Birth_Data_1b)[names(df_Birth_Data_1b) == "b4_03"] <- "Sexbirth3"


df_Birth_Data_1b
names(df_Birth_Data_1b)

# Convert wide dataframe into long dataframe - Birthyear
df_Birth_Data_1b_long <- gather(df_Birth_Data_1b,BirthyearCategory,Birthyear,BirthYear1,Birthyear2,Birthyear3) 
# %>% arrange(MothersAge)
df_Birth_Data_1b_long

# Convert wide dataframe into long dataframe - SexBirth
df_Birth_Data_1c_long <- gather(df_Birth_Data_1b_long,SexBirthCategory,Gender,Sexbirth1,Sexbirth2,Sexbirth3) 
# %>% arrange(MothersAge)
df_Birth_Data_1c_long

df_Birth_Data_1d_long <- df_Birth_Data_1c_long
df_Birth_Data_1d_long
# Cleanup dataframe with appropriate values
df_Birth_Data_1d_long$Gender <- gsub("1", "M", df_Birth_Data_1d_long$Gender)
df_Birth_Data_1d_long$Gender <- gsub("2", "F", df_Birth_Data_1d_long$Gender)
df_Birth_Data_1d_long$Gender <- gsub("NA", "", df_Birth_Data_1d_long$Gender)
df_Birth_Data_1d_long$Birthyear <- gsub("NA", "", df_Birth_Data_1d_long$Birthyear)

df_Birth_Data_1d_long

# Plot-1 Range of values per Mothers Age and BirthyearCategory
qplot(x=MothersAge,y=BirthyearCategory, data=df_Birth_Data_1d_long, geom="line",colour = Gender)


# Plot-2 range of MothersAge values per BirthyearCategory
qplot(x=MothersAge,y=Gender, data=df_Birth_Data_1d_long, geom="line") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# Plot-3 range of Percent_Income values per Religious.tradition by IncomeRange
ggplot(df_Birth_Data_1d_long, aes(BirthyearCategory,Gender, group = MothersAge, colour = MothersAge)) + geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))



```
   

Conclusion: In Plot-2 it can be seen that more of the younger mothers had a Male child.  In Plot-3 it can be seen that there are instances of a first child male and second child female, a first child female and second child male, and first child male and second child male, as well as first child female and second child female.





#END
