---
title: "R_607_Project_2_Wide_to_Long_Transform_Data_Daniel_Thonn"
output: html_document
course: R-607 CUNY MSDA program

---

## Assignment 607_Project_2_Wide_to_Long_Transform_Concrete_Blocks_Daniel_Thonn

Summary of Assignment
This assignment involves Tidying and Transforming data

This Assignment requires the following:

1). R-Studio

The following R-packages are used:
1.tidyr
2.dplyr
3.ggplot2 

Steps to reproduce:
1). Place file on local system: C:/mysqldata/Construction_Materials_Table_Wide_1d4.csv
2). run the R-Studio file: R_607_Project_2_Wide_to_Long_Daniel_Thonn_1b.Rmd



Setting up and Preparing the Environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}

#install.packages("stringr")
#install.packages("tidyr")
#install.packages("dplyr")
#install.packages("plyr")
#library(stringr)
library(tidyr)
library(dplyr)
#library(plyr)
#detach("package:plyr", unload=TRUE)
library(ggplot2)

```


Load the character names data into an R dataframe, cleanup, and convert to long format


```{r, echo=TRUE}

# Load the data csv file to a vector
b <- read.csv(
              "C:/mysqldata/Construction_Materials_Table_Wide_1d4.csv",
              sep=",",
              na.strings = "",
              blank.lines.skip = TRUE,
                stringsAsFactors=FALSE)
list(b)

# Convert the input vector to a dataframe
df_Const_Matls_1 = data.frame(b)
list(df_Const_Matls_1)
df_Const_Matls_1b <- df_Const_Matls_1
#names(df_Const_Matls_1b)[names(df_Const_Matls_1b) == "X2009"] <- "2009"
#names(df_Const_Matls_1b)[names(df_Const_Matls_1b) == "X2010"] <- "2010"
#names(df_Const_Matls_1b)[names(df_Const_Matls_1b) == "X2011"] <- "2011"
#names(df_Const_Matls_1b)[names(df_Const_Matls_1b) == "X2012"] <- "2012"
#names(df_Const_Matls_1b)[names(df_Const_Matls_1b) == "X2013"] <- "2013"
#names(df_Const_Matls_1b)


# Convert wide dataframe into long dataframe
df_Const_Matls_1_long <- gather(df_Const_Matls_1,Year,Price_Index, X2009:X2013) %>% arrange(Material.Category)
#df_Const_Matls_1_long <- gather(df_Const_Matls_1b,Year,Price_Index, 2009:2013) %>% arrange(Material.Category)

list(df_Const_Matls_1_long)
df_Const_Matls_1a_long <- df_Const_Matls_1_long
df_Const_Matls_1a_long$Year <- gsub("X", "",df_Const_Matls_1_long$Year)
list(df_Const_Matls_1a_long)

df_Const_Matls_1b_long <- df_Const_Matls_1a_long

# Remove spaces from column Price_Index values
df_Const_Matls_1_long$Price_Index
df_Const_Matls_1b_long$Price_Index <- as.numeric(gsub("\\..{2}", "0",df_Const_Matls_1_long$Price_Index))
df_Const_Matls_1b_long$Price_Index


# Remove NA's in Price_Index column
df_Const_Matls_1c_long <- df_Const_Matls_1b_long
df_Const_Matls_1c_long$Price_Index <- replace(df_Const_Matls_1b_long$Price_Index, is.na(df_Const_Matls_1b_long$Price_Index), 0) 

df_Const_Matls_1c_long$Price_Index
names(df_Const_Matls_1c_long)

# Test
#p <- ggplot(df_Const_Matls_1c_long, aes(x=Year, y=Price_Index, group=Materials.Type)) + geom_line()
#p 

# Plot Range of values per year
qplot(x=Year,y=Price_Index, data=df_Const_Matls_1c_long, geom="line")


# Plot range of price_index values per Materials.Type
qplot(x=Materials.Type,y=Price_Index, data=df_Const_Matls_1c_long, geom="line") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# Plot range of price_index values per Materials.Type per year
ggplot(df_Const_Matls_1c_long, aes(Materials.Type, Price_Index, group = Year, colour = Year)) +
  geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


# Setup group by criteria by Materials.Type
By_Materials_Type_1 <- group_by(df_Const_Matls_1c_long,Materials.Type,Year) 
By_Materials_Type_1

By_Materials_Type_1 %>% ggplot(aes(x=Year, y=Price_Index)) + geom_line()


# Select columns for analysis dataframe by Materials.Type
By_Materials_Type_2 <- select(By_Materials_Type_1,Materials.Type,Year,Price_Index)
By_Materials_Type_2


# Present columns based on select and group-by criteria
By_Materials_Type_3 <- summarise(By_Materials_Type_2,Mean_Price_Index = mean(Price_Index))
By_Materials_Type_3

# Graph Materials.Type and Price_Index per Year
By_Materials_Type_3 %>%
ggplot(aes(Materials.Type, Mean_Price_Index, group = Year, colour = Year)) +
  geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

# Graph by Mean_Price_Index per year
By_Materials_Type_3 %>%
ggplot(aes(Year, y = Mean_Price_Index,group = Year)) +
  geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```
   

Conclusion: 
The overall Mean_Price_Index dipped in 2010, but has been steadily increasing from 2010 to 2014.










#END
