---
title: "R_607_Project_2_Wide_to_Long_Transform_Data_Daniel_Thonn"
output: html_document
course: R-607 CUNY MSDA program

---

## Assignment 607_Project_2_Wide_to_Long_Transform_Income_Distribution_By_Religion_Daniel_Thonn

Summary of Assignment
This assignment involves Tidying and Transforming flight data

This Assignment requires the following:

1). R-Studio

The following R-packages are used:
1.tidyr
2.dplyr
3.ggplot2 

Steps to reproduce:
1). Place file locally: C:/mysqldata/Income_Distribution_by_Religion.csv

2). run the R-Studio file: R_607_Project_2_Wide_to_Long_Daniel_Thonn_2a.Rmd


Setting up and Preparing the Environment

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}

#install.packages("stringr")
#install.packages("tidyr")
#install.packages("dplyr")
#install.packages("plyr")
#library(stringr)
library(tidyr)
library(dplyr)
#library(plyr)
#detach("package:plyr", unload=TRUE)
library(ggplot2)

```


Load the character names data into an R dataframe, cleanup, and convert to long format


```{r, echo=TRUE}

# Load the data csv file to a vector
c <- read.csv(
              "C:/mysqldata/Income_Distribution_by_Religion.csv",
              sep=",",
              na.strings = "",
              blank.lines.skip = TRUE,
                stringsAsFactors=FALSE)
list(c)

# Convert the input vector to a dataframe
df_Income_Religion_1 = data.frame(c)
df_Income_Religion_1
df_Income_Religion_1b = df_Income_Religion_1[, !(colnames(df_Income_Religion_1) %in% c("Sample.Size"))]
df_Income_Religion_1b

names(df_Income_Religion_1b)[names(df_Income_Religion_1b) == "Less.than..30.000"] <- "Lessthan30"
names(df_Income_Religion_1b)[names(df_Income_Religion_1b) == "X.30.000....49.999"] <- "In30to49"
names(df_Income_Religion_1b)[names(df_Income_Religion_1b) == "X.50.000....99.000"] <- "In50to99"
names(df_Income_Religion_1b)[names(df_Income_Religion_1b) == "X.100.000.or.more"] <- "In100plus"

df_Income_Religion_1b

names(df_Income_Religion_1b)


# Convert wide dataframe into long dataframe
df_Income_Religion_1_long <- gather(df_Income_Religion_1b,IncomeRange,Percent_Income,Lessthan30,In30to49,In50to99,In100plus) %>% arrange(Religious.tradition)
df_Income_Religion_1_long


# Plot-1 Range of values per IncomeRange
qplot(x=IncomeRange,y=Percent_Income, data=df_Income_Religion_1_long, geom="line")


# Plot-2 range of Percent_Income values per Religious.tradition
qplot(x=Religious.tradition,y=Percent_Income, data=df_Income_Religion_1_long, geom="line") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

# Plot-3 range of Percent_Income values per Religious.tradition by IncomeRange
ggplot(df_Income_Religion_1_long, aes(Religious.tradition, Percent_Income, group = IncomeRange, colour = IncomeRange)) + geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


# Setup group by criteria by IncomeRange
By_Income_Range_1 <- group_by(df_Income_Religion_1_long,Religious.tradition,IncomeRange) 
By_Income_Range_1

By_Income_Range_1 %>% ggplot(aes(x=IncomeRange, y=Percent_Income)) + geom_line()


# Select columns for analysis dataframe by IncomeRange
By_Income_Range_2 <- select(By_Income_Range_1,Religious.tradition,IncomeRange,Percent_Income)
By_Income_Range_2


# Present columns based on select and group-by criteria
By_Income_Range_3 <- summarise(By_Income_Range_2,Mean_Percent_Income = mean(Percent_Income))
By_Income_Range_3

# Plot-4: Graph Religious.tradition and Mean_Percent_Income per IncomeRange
By_Income_Range_3 %>%
ggplot(aes(Religious.tradition, Mean_Percent_Income, group = IncomeRange, colour = IncomeRange)) +
  geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

# Plot-5: Graph by Mean_Percent_Income per Income Range 
By_Income_Range_3%>%
ggplot(aes(IncomeRange, y = Mean_Percent_Income,group = Religious.tradition, colour = Religious.tradition)) +
  geom_path(alpha = 0.5) + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```
   

Conclusion: The Percent_Income range is widest in the Lessthan30K and In100plus range income.  Also as shown in Plot-2 the widest range of Percent_Income is widest among Historically Black Protestant and Jehovahs witness and narrowest among Mainline Protestant and Mormon. Plot-4 shows that the Mean_Percent_Income is lowest among Historically Black Protestant and highest among Jewish.   Plot-5 shows that the widest disparities by religions is in the 100k plus range and in the Lessthan30k range.





#END
